---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Riccardo Fusaroli"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia, focusing on pitch.
In the course of this assignment we will use them to try to automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.

### Question 1: Can you diagnose schizophrenia from pitch range only? If so, how well?

Build a logistic regression to see whether you can diagnose schizophrenia from pitch range only.

Calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve) on a logistic regression using the full dataset. Don't forget the random effects!

Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures.

N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?

```{r}

# To load libraries
library(lmerTest); library(caret);library(pROC); library(cvms); library(groupdata2); library(magrittr); library(dplyr)

# CVMS is a possible solution - Made from last year students

# To set working directory
setwd("C:/Users/simon/Google Drev/Uni/Methods3/Assignment3")

# To load file
S_data = read.csv("Schizophrenia_data.csv")
S_data$Subject = as.factor(S_data$Subject)

# To deleta anoying columns
S_data=S_data[,-1:-2]

# To create a logistic regression
m1=glmer(Diagnosis ~ range + (1+Trial|Subject), S_data, family = "binomial")
summary(m1)

# To create a confusion matrix
S_data$PredictionsPerc=predict(m1) 
S_data$Predictions[S_data$PredictionsPerc>0]="Schizophrenia" 
S_data$Predictions[S_data$PredictionsPerc<=0]="Control"

confusionMatrix(data = S_data$Predictions, reference = S_data$Diagnosis, positive = "Schizophrenia") 

# To create the ROC plot
rocCurve <- roc(response = S_data$Diagnosis,   predictor = S_data$PredictionsPerc) 
auc(rocCurve) 
ci(rocCurve) 
plot(rocCurve, legacy.axes = TRUE) 


# To add fold_value
fold_function = function(data, col) {

  data$folds_ID = as.numeric(interaction(data[,col]))
  
  return(data[order(data$folds_ID),])
}

Data_Schizo = subset(S_data, S_data$Diagnosis == "Schizophrenia")
Data_Control = subset(S_data, S_data$Diagnosis == "Control")

Data_Schizo=fold_function(Data_Schizo, "Subject")
Data_Control= fold_function(Data_Control, "Subject")

# To create folds
folds_S = createFolds(unique(Data_Schizo$folds_ID), 5)
folds_C= createFolds(unique(Data_Control$folds_ID), 5)

# To combine data
fold_list = c()
fold_list$Fold1 = c(folds_S$Fold1, folds_C$Fold1)
fold_list$Fold2 = c(folds_S$Fold2, folds_C$Fold2)
fold_list$Fold3 = c(folds_S$Fold3, folds_C$Fold3)
fold_list$Fold4 = c(folds_S$Fold4, folds_C$Fold4)
fold_list$Fold5 = c(folds_S$Fold5, folds_C$Fold5)

NewDATA = rbind(Data_Schizo, Data_Control)

# To create empty variables
Accuracy = NULL
Specificity = NULL
Sensitivity = NULL
PPV = NULL
NPV = NULL
AUC = NULL

n = 1

# To create a loop
for (fold in fold_list){
  
  # Create training data set
  train = subset(NewDATA,! (folds_ID %in% fold))  
  
  # Create a test data set
  test = subset(NewDATA, (folds_ID %in% fold)) 
  
  # To train model
  m1=glmer(Diagnosis ~ range + (1+Trial|Subject), train, family = "binomial")
  
  # To evaluate model
  test$PredictionsPerc=predict(m1, test, allow.new.levels = TRUE) 
  test$Predictions[test$PredictionsPerc>0]="Schizophrenia" 
  test$Predictions[test$PredictionsPerc<=0]="Control"
  
  cm=confusionMatrix(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia") 
  
  Accuracy[n] = cm$overall["Accuracy"]
  
  test$Predictions = as.factor(test$Predictions)

  Sensitivity[n] =  sensitivity(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia" )
  
  Specificity [n] = specificity(data = test$Predictions, reference = test$Diagnosis, negative = "Control" ) 

  PPV[n] = posPredValue(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia") 

  NPV[n] = negPredValue(data = test$Predictions, reference = test$Diagnosis, negative = "Control")  
  
  rocCurve = roc(response = test$Diagnosis,   predictor = test$PredictionsPerc) 
  AUC[n]=auc(rocCurve) 
    
  n=n+1  
}

# To create the dataframe
test_data = data.frame(Accuracy, Sensitivity, Specificity, PPV, NPV, AUC)

# To get the mean AUC
mean(test_data$AUC) #0.59


```


### Question 2 - Which single acoustic predictor is the best predictor of diagnosis?

```{r}

# To create a function to find out which single acoustic predictor is the best predictor
CV_function = function(data, model) {

# To create empty variables
Accuracy = NULL
Specificity = NULL
Sensitivity = NULL
PPV = NULL
NPV = NULL
AUC = NULL

n = 1

# To create a loop
for (fold in fold_list){
  
  # Create training data set
  train = subset(data,! (folds_ID %in% fold))  
  
  # Create a test data set
  test = subset(data, (folds_ID %in% fold)) 
  
  # To train model
  m1=glmer(model, train, family = "binomial")
  
  # To evaluate model
  test$PredictionsPerc=predict(m1, test, allow.new.levels = TRUE) 
  test$Predictions[test$PredictionsPerc>0]="Schizophrenia" 
  test$Predictions[test$PredictionsPerc<=0]="Control"
  
  cm=confusionMatrix(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia") 
  
  Accuracy[n] = cm$overall["Accuracy"]
  
  test$Predictions = as.factor(test$Predictions)

  Sensitivity[n] =  sensitivity(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia" )
  
  Specificity [n] = specificity(data = test$Predictions, reference = test$Diagnosis, negative = "Control" ) 

  PPV[n] = posPredValue(data = test$Predictions, reference = test$Diagnosis, positive = "Schizophrenia") 

  NPV[n] = negPredValue(data = test$Predictions, reference = test$Diagnosis, negative = "Control")  
  
  rocCurve = roc(response = test$Diagnosis,   predictor = test$PredictionsPerc) 
  AUC[n]=auc(rocCurve) 
    
  n=n+1  
}

results = data.frame(Accuracy, Sensitivity, Specificity, PPV, NPV, AUC)

return(results)

}

mean_results = CV_function(NewDATA, Diagnosis ~ mean + (1+Trial|Subject))
SD_results = CV_function(NewDATA, Diagnosis ~ SD + (1+Trial|Subject))
median_results = CV_function(NewDATA, Diagnosis ~ median + (1+Trial|Subject))
RR_results = CV_function(NewDATA, Diagnosis ~ RR + (1+Trial|Subject))
ENTR_results = CV_function(NewDATA, Diagnosis ~ ENTR + (1+Trial|Subject))
rENTR_results = CV_function(NewDATA, Diagnosis ~ rENTR + (1+Trial|Subject))
IQR_results = CV_function(NewDATA, Diagnosis ~ IQR + (1 + Trial|Subject))
MAD_reults = CV_function(NewDATA, Diagnosis ~ MAD + (1 + Trial|Subject))
L_results = CV_function(NewDATA, Diagnosis ~ L + (1 + Trial|Subject))
Lmax_results = CV_function(NewDATA, Diagnosis ~ maxL + (1 + Trial|Subject))

# To obtain the mean area under the curve, which is a measure of overall performance
mean(SD_results$AUC) # 0.6143
mean(mean_results$AUC) # 0.57
mean(median_results$AUC) # 0.57
mean(RR_results$AUC) # 0.51
mean(ENTR_results$AUC) # 0.615
mean(rENTR_results$AUC) # 0.56
mean(IQR_results$AUC) # 0.6116
mean(MAD_reults$AUC) #0.6146
mean(L_results$AUC) # 0.588
mean(Lmax_results$AUC) # 0.53


```

### Question 3 - Which combination of acoustic predictors is best for diagnosing schizophrenia?

Now it's time to go wild! Use all (voice-related) variables and interactions you can think of. Compare models and select the best performing model you can find.

Remember:
- Out-of-sample error crucial to build the best model!
- After choosing the model, send Celine and Riccardo the code of your model

```{r}
# To create different models with combination of features 
comb1_results = CV_function(NewDATA, Diagnosis ~ range*mean + (1+Trial|Subject))
comb2_results = CV_function(NewDATA, Diagnosis ~ range*mean + L + (1+Trial|Subject))
comb3_results = CV_function(NewDATA, Diagnosis ~ range*L + (1+Trial|Subject))
comb4_results = CV_function(NewDATA, Diagnosis ~ SD*ENTR + (1+Trial|Subject))
comb5_results = CV_function(NewDATA, Diagnosis ~ SD+ENTR + (1+Trial|Subject))

# To get the mean area under the curve for each model 
mean(comb1_results$AUC) # 0.62
mean(comb1_results$Accuracy)
mean(comb2_results$AUC) # 0.618
mean(comb3_results$AUC) # 0.61
mean(comb4_results$AUC) # 0.61
mean(comb5_results$AUC) # 0.617

```

### Question 4: Properly report the results

METHODS SECTION: how did you analyse the data? That is, how did you extract the data, designed the models and compared their performance?

RESULTS SECTION: can you diagnose schizophrenia based on voice? which features are used? Comment on the difference between the different performance measures.

### Bonus question 5

You have some additional bonus data involving speech rate, pauses, etc. Include them in your analysis. Do they improve classification?

### Bonus question 6

Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them.
